\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-09-16 chapters/bci.tex
Add table, references for eye motor disabilities

@@

Simply put, a BCI records the user's brain activity, then extracts some
relevant output from this brain activity and couples this output to a function
of a device.
Optionally, the user can then observe the action of the device and adapt their
behaviour or brain activity accordingly, closing the loop
\footnote{
  \textcite{BCISociety2024} has recently formalized this into the
  following definition of a BCI:
  \it``A brain-computer interface is a system that measures brain activity and
  converts it in (nearly) real-time into functionally useful outputs to replace,
  restore, enhance, supplement, and/or improve the natural outputs of the brain,
  thereby changing the ongoing interactions between the brain and its external or
  internal environments. It may additionally modify brain activity using targeted
  delivery of stimuli to create functionally useful inputs to the
  brain.''
}.

@@

Let's break this definition down and focus on its separate parts.
First of all, we need to identify a signal that is a direct representation of
what's going on in the brain.
One might for instance measure the fluctuations in blood flow to specific brain
regions that are more or less active at a given time, i.e. the
\textit{haemodynamic signal} with a scanner.
However, this signal reacts to slowly to reflect the real-time activity that is
of interest for our purposes, and it carries little information other than the
brain regions where it is originating.
A better candidate is the neuronal \textit{action potential}.
The brain consists 86 billion neurons, which are highly interconnected cells that are
the smallest units of information processing.
The action potential is an electrical pulse occurring when a neuron receives
input from another neuron and is activated.\todo{kort door de bocht}
A neuron's action potential, or the combined action potentials of groups of
synchronized neurons, thus generate an electrical field in and near the brain,
which can be measured using electrodes.
Fluctuations in the brain's electrical potential are a much faster and more
informative signal for our purposes.
\todo{lower part to section}
While a myriad of other techniques can be used to measure functional brain
activity, such as functional ultrasound imaging (fUS) or magnetoencephalograpy
(MEG), we'll limit ourselves in this work to measuring the electrical field.
Section~\ref{sec:bci-recording} gives an overview of current recording
technologies.\todo{make shorter, move part to section}

\todo{paradigm}
The brain waves form only one part of the interaction scheme presented
in Figure~\ref{fig:bci-loop}.
Most importantly, wee need to know what exactly to look for in this signal.
For this, the user needs to perform a task which evokes specific activity of
which it is known that it can be retrieved from the brain signal.
For example, the imagined movement of an arm is detectable f

Next, this brain signal needs to be interpreted \emph{in function of} the
task we want our BCI to perform.\todo{feature extraction}
The electrical activity of the brain is relatively weak compared to that of the
environment around us, containing electronic devices and other sources of
interference.
Furthermore, the brain is continuously processing information and carrying out
`background tasks', all of which generate their own electrical signals.
On the other hand, the brain activity we are interested in exploiting often
originates from a specific functional system within the brain, and its activity
might not be easily discerned from all other neural and environmental activity,
akin to trying to listen to a single speaker in a noise environment where
everyone is shouting over each other.
Some obvious sources of interference can be easily filtered out, but to obtain
a good measure reflecting only the activity that is relevant for a given BCI
problem.
This problem is called \emph{decoding} and is often solved through supervised
machine learning.
Section~\label{sec:bci-decoding} takes a closer look at some common and
state-of-the art techniques.
A machine learning algorithm can use some training data for which the performed
tasks are known, and learn to recognize these tasks in new data.
The training data can be either obtained from the BCI user themselves or from
other users.
In the first case, the user would perform a short calibration session before
using the BCI, where they are instructed to perform a known task.
This calibration session can be eliminated by training the decoder on
preexisting labeled data from other sessions and users, but this is harder due
to the variability between subjects and sessions.\todo{make shorter, move part
to section}

@@

@@

\section{Recording technologies}
\label{sec:bci-recording}
different recording technologies and level of invasiveness + infographic
cost, ease of use
\section{BCI Paradigms}
\label{sec:bci-paradigms}
\fullpagefig{figures/bci/paradigms.eps}{caption}{fig:bci-paradigms}

>>>>

tetraplegic, locked-in syndrome
Regaining capabilities and improving QoL
Start from wolpaw examples, for spelling
Cite State of the art examples (walk again, speech avatar Metzger, handwriting
willet, neuralink)

@@

Simply put, a BCI records the user's brain activity, then extracts some
relevant output from this brain activity and couples this output to a function
of a device.
Optionally, the user can then observe the action of the device and adapt their
behaviour or brain activity accordingly, closing the loop
\footnote{
  \textcite{BCISociety2024} has recently formalized this into the
  following definition:
  \it``A brain-computer interface is a system that measures brain activity and
  converts it in (nearly) real-time into functionally useful outputs to replace,
  restore, enhance, supplement, and/or improve the natural outputs of the brain,
  thereby changing the ongoing interactions between the brain and its external or
  internal environments. It may additionally modify brain activity using targeted
  delivery of stimuli to create functionally useful inputs to the
  brain.''
}.

@@

Let's break this definition down and focus on its separate parts.
First of all, we need to identify a signal that is a direct representation of
what's going on in the brain.
One might for instance measure the fluctuations in blood flow to specific brain
regions that are more or less active at a given time, i.e. the
\emph{hemodynamic signal} with a scanner.
However, this signal reacts to slowly to reflect the real-time activity that is
of interest for our purposes, and it carries little information other than the
brain regions where it is originating.
A better candidate is the neuronal \emph{action potential}.
The brain consists 86 billion neurons, which are highly interconnected cells that are
the smallest units of information processing.
The action potential is an electrical pulse occurring when a neuron receives
input from another neuron and is activated.\todo{kort door de bocht}
A neuron's action potential, or the combined action potentials of groups of
synchronized neurons, thus generate an electrical field in and near the brain,
which can be measured using electrodes.
Fluctuations in the brain's electrical potential can be faster and more
informative.\todo{ref section}

The recorded brain waves form only one part of the interaction scheme presented
in Figure~\ref{fig:bci-loop}.
It would make little sense to start digging around at random in all the brain
activity to try to extract the desired output.
It would help greatly if we knew what we are looking for.
For this reason, a BCI operates using a defined \emph{paradigm}, the manner in
which interaction is performed.
This usually means that the user is instructed to perform a specific task, like
attending a specific flickering stimulus.
We can now exploit background knowledge from prior neuroscientific research
that the brain response evoked by attended stimuli is different of those
unattended,
and determine which one of a set of targets was attended.
In turn, this can be coupled to an action, like typing the letter shown on the
stimulus.
The BCI paradigm comprises, if applicable, the manner of stimulation and the
exact task performed by the user, and is often closely linked to the manner of
feedback in the case of a closed-loop system.
As shown in Section~\ref{sec:bci-paradigms}, there is a wide variety of BCI
paradigms based on different systems in the brain.

The previous points to another component of a successful BCI: the specific
brain signals related to our paradigm need to be identified within the recorded
activity.
This is commonly referred to as \emph{feature extraction}
In other words, the signal needs to be interpreted \emph{in function of} the
task we want our BCI to perform.
The electrical activity of the brain is relatively weak compared to that of the
environment around us, containing electronic devices and other sources of
interference.
Furthermore, the brain is continuously processing information and carrying out
`background tasks', all of which generate their own electrical signals.
On the other hand, the brain activity we are interested in exploiting often
originates from a specific functional system within the brain, and its activity
might not be easily discerned from all other neural and environmental activity,
akin to trying to listen to a single speaker in a noise environment where
everyone is shouting over each other.
Some obvious sources of interference can be easily filtered out, but to obtain
a good measure reflecting only the activity that is relevant for a given BCI
problem.
This problem is called \emph{decoding} and is often solved through supervised
machine learning.
Section~\label{sec:bci-decoding} takes a closer look at some common and
state-of-the art techniques.
A machine learning algorithm can use some training data for which the performed
tasks are known, and learn to recognize these tasks in new data.
The training data can be either obtained from the BCI user themselves or from
other users.
In the first case, the user would perform a short calibration session before
using the BCI, where they are instructed to perform a known task.
This calibration session can be eliminated by training the decoder on
preexisting labeled data from other sessions and users, but this is harder due
to the variability between subjects and sessions.\todo{make shorter, move part
to section}

@@

\todo{wrap-up}

@@

\section{Recording technologies}
\label{sec:bci-recording}
The brain's activity can be recorded using various neuroimaging
technologies.
These can range from brain scans~\cite{Weiskopf2004} using functional magnetic resonance imaging
(fMRI) to more portable technologies like accoustic signals obtained by functional ultrasound imaging
(fUS)~\cite{Zheng2023} and near-infrared spectroscopy
(fNIRS)~\cite{Borgheai2020}.
These technologies all measure the hemodynamic signal, which only indirectly
reflects brain activity.
Because of the desirable high temporal resolution of
electrophysiology~\cite{Easttom2021} and
because of their practicality, we will focus only on methods to record
electromagnetical activity originating from neuronal action potentials.
The magnetical field of the brain can be measured using magnetoencephalography
(MEG)~\cite{Mellinger2007}.
While MEG is non-invasive and results in high-quality signals, it is relatively
impractical due to the size of the equipment needed.
Recent advances have been made using optically pumped magnetometers
(OPM-MEG)~\cite{Wittevrongel2021}, but this technology still falters outside of
lab conditions.

As a consequence, the BCI field relies heavily on technologies measuring the
electrical activity of the brain.
As mentioned earlier with MEG, an important concept in determining which technology within this group is
suited for a BCI application is the invasiveness of the technology.
single-unit recordings
microelectrode arrays (utah-array, neuralink)
ecog/µecog
eeg

EEG cheap and practical, non-invasive, high-temporal resolution, but noise, suffers from volume
conduction hence low spatial resolution
what does EEG measure?

different recording technologies and level of invasiveness + infographic
cost, ease of use
\section{BCI Paradigms}
\label{sec:bci-paradigms}
\fullpagefig{figures/bci/paradigms.eps}{caption}{fig:bci-paradigms}

### 2024-09-16 chapters/gaze_independence.tex
Add table, references for eye motor disabilities

One of the goals of Brain-computer interfacing is establishing a communication
channel that does not rely on speech or
muscular activity~\cite{Naci2012,Chaudhary2016}, which in turn can provide
solutions to paralyzed individuals.
%The quest for performant and affordable solutions is most evident in
%visual BCIs using the electroencephalogram (EEG), where it has led to a gamut
%of increasingly sophisticated decoders, tailored to the needs
%of specific stimulation paradigms and usage contexts.

@@

In this work, we will use the term \emph{gaze-independent} meaning
\emph{‘dealing explicitly with the fact that a user cannot control their
gaze.'}
In the context of a visual BCI, this means that the user's
visuospatial attention and their gaze do not necessarily coincide.
Gaze-independent paradigms are particularly
promising for individuals with impaired eye motor control, such as those
suffering from specific types or stages of Amyotrophic Lateral Sclerosis (ALS),
Multiple Sclerosis (MS), stroke, or brain stem
stroke.
For these patients, gazing directly at a screen target may be uncomfortable, impractical,
or even impossible.
Hence, assistive devices that rely on eye tracking are often inefficient for
them.
Consequently, while BCIs hold great promise for these individuals, conventional
gaze-dependent BCI solutions do not meet their needs due to the absence of gaze
control. Therefore, the development of decoding strategies that account for covert
VSA becomes crucial in the pursuit of high-performance gaze independent
BCIs.

\section{Incidence of eye motor impairment}
Table~\ref{tab:incidence} reports the relatively high frequencies of
eye motor problems, which can range from involuntary eye movements to control
issues and even to partial or complete ophtalmoplegia.
\begin{table}
	\centering
	\begin{tabular}{@{}|l|rrrrr|r|@{}}
		\hline
		         & ALS  & MS   & Stroke  &
		DMD      & SMA  & LiS                           \\ \hline
		Minor    & 50\% & 31\% & 40-70\% & + & - &      \\
		Severe   & 33\% & 3\%  & +       & - & - & 98\% \\
		Complete & 17\% & -    & +       & - & - & 2\%  \\
		\hline
	\end{tabular}
	\caption{Incidence of eye motor impairment in Brain-Computer Interface target
		populations: (ALS: Amyotrophic Lateral Sclerosis, MS: Multiple Sclerosis,
		DMD: Duchenne's Muscular Dystrophy, SMA: Spinal Muscular Atrophy, LiS:
		Locked-In-Syndrome).}\label{tab:incidence}
\end{table}
\todo{document methodology}
\todo{find more references, fill out}
\todo{
Patients with
Totally locked in/locked in with eye motor issues due to
Brain stem stroke
ALS
MS
Motor and eye motor impairments due to neurotrauma (tetraplegia)
...
}

@@

>>>>

One of the goals of Brain-computer interfacing is establishing a communication
channel that does not rely on speech or
muscular activity~\cite{Naci2012,Chaudhary2016}, which in turn can provide
solutions to paralyzed individuals.
In the strictest interpretation, this also means that an interface should not
rely on the control of eye muscles which allow the subject to redirect their
gaze or stare at a target.

In fact, many patients in the BCI target population can suffer to some extent
from eye motor disability, requiring BCIs adapted to their condition.
Table~\ref{tab:incidence} reports the relatively high frequencies of
eye motor problems, which can range from minor (nystagmus, other tremors, gaze
fatigue or discomfort, \ldots), to severe (partial ophtalmoplegia, involuntary
movements, impaired pursuit, \ldots) and even complete ophtalmoplegia.
issues and even to partial or complete ophtalmoplegia.
\begin{table}[h]
  \sffamily
  \footnotesize
	\centering
	\begin{tabular}{@{}l|rrrrr|r@{}}
              & \bfseries ALS  & \bfseries MS   & \bfseries Stroke  &\bfseries DMD &\bfseries SMA  &\bfseries LiS  \\ \hline
		\bfseries Minor    & 50\% & 31\% & 40-70\% & + & - &      \\
		\bfseries Severe   & 33\% & 3\%  & +       & - & - & 98\% \\
		\bfseries Complete & 17\% & -    & +       & - & - & 2\%  \\
	\end{tabular}
	\caption{Incidence of eye motor impairment in selected BCI target
		patient populations. (ALS: Amyotrophic Lateral Sclerosis, MS: Multiple Sclerosis,
		DMD: Duchenne's Muscular Dystrophy, SMA: Spinal Muscular Atrophy, LiS:
		Locked-In-Syndrome).}\label{tab:incidence}
\end{table}
Among the most affected patients are ALS patients, especially those suffering
from bulbar onset ALS~\cite{Guo2022}.
Stroke, brain-stem and cerebellar stroke.
Furthermore, patients with Locked-in Syndrome
\footnote{Multiple definitions of Locked-in Syndrome are encountered in
BCI and neurological literature. Some definitions include only
tetraplegic patients without eye movements used for communications. Others distinguish Complete
Locked-in Syndrome (CLIS) with full body paralysis, including no eye motor
control at all, from a LiS state with some preserved eye movements or minor
motor output.
While some definitions only include patients who suffered stroke or traumatic
brain injury to the brain stem, it can also generally refer to the state of full body paralysis
or loss of muscle tone, combined with the inability to speak, such as occurs in
late stage ALS.
In clinical reality, every tetraplegic patient has their
unique set of preserved capabilities.}
\todo{fill out which definition we use, wolpaw criteria}
\todo{cite studies per pathology}
Gaze-independent paradigms are particularly
promising for individuals with impaired eye motor control, such as those
suffering from specific types or stages of Amyotrophic Lateral Sclerosis (ALS),
Multiple Sclerosis (MS), (brain stem) stroke or Friedreich's Ataxia (FA)

@@

In this work, we will use the term \emph{gaze-independent} meaning
\emph{‘dealing explicitly with the fact that a user cannot control their
gaze.'}
In the context of a visual BCI, this means that the user's
visuospatial attention and their gaze do not necessarily coincide.
For patients with eye motor impairment, gazing directly at a screen target may
be uncomfortable, impractical, or even impossible.
Hence, assistive devices that rely on eye tracking are often inefficient for
them.
Consequently, while BCIs hold great promise for these individuals, conventional
gaze-dependent BCI solutions do not meet their needs due to the absence of gaze
control. Therefore, the development of decoding strategies that account for covert
VSA becomes crucial in the pursuit of high-performance gaze independent
BCIs.

@@

\todo{provisionary doctoral plan}
\todo{check WIP seminar}
\todo{update text}

\section{Problem statement}

\subsection{Objectives}

Brain-Computer Interfaces (BCIs) decode brain activity with the aim to establish a direct communication
pathway bypassing speech and other forms of muscular activity. They have raised great hopes for patients
devoid of these abilities, for whom BCIs can provide a means to communicate or to
control devices.
The quest for performant and affordable solutions is most evident in
visual BCIs based on the electroencephalogram (EEG), where it has led to a gamut of increasingly sophisticated decoders
and paradigms tailored to the needs of specific stimulation paradigms and
use contexts.
An effective and proven method is the visual event-related potential (ERP)-based
oddball interface.
Here, targets are shown in short pulses on a computer screen.
Each time the user observes a target, an ERP is evoked, the presence of which can be
detected in the EEG-signal.
The ERP consists of multiple components, of which some are modulated by the
attention of the user, specifically the N2 and the P3 component.
Decoding this modulation allows for information transfer controlled by the
user's brain activity, as sketched in Figure~\ref{fig:bci}.
However, visual BCIs cannot operate efficiently when the user does not direct
their gaze onto the desired target~\cite{Brunner2010, Frenzel2011}.

This leads to an often unadressed paradox: the BCI target population consists
of patients who suffer from pathologies like Amyotrophic Lateral Sclerosis
(ALS), Multiple Sclerosis (MS), stroke (specifically
brain-stem stroke) or traumatic brain injury, which can lead to several degrees
of paralysis, or even to Locked-in Syndrome (LiS), the complete loss of muscle
control with preserved consciousness.
While BCIs are most attractive as a solution for these patients
compared to other assistive technologies like eye-tracking, studies often
report bad performance in patient populations precisely due to the lack of adequate eye motor
control, or patients with gaze impairments are excluded.
Different degrees of eye motor impairment can render visual BCIs
uncomfortable or outright impossible to use.
When operating a visual BCI, the usual rapid series of forced saccades followed
by fixation is tiring over time, even for healthy control subjects.
A suitable alternative would allow the user to keep their eyes in an at-rest
position of their choice while operating the BCI.
This points to the need for gaze-independent BCIs.

The general goal of this PhD is to tackle the gaze-dependence in visual BCIs.
\todo{refer to ITRs from previous chapter}
Section{sec:gaze-independence-paradigms} showed that visual BCI paradigms form
a good candidate for a performant gaze-independent BCI.
While the central gaze fixation of most visual gaze-independent paradigms still
relies to some extent on eye motor control, we aim to circumvent this.
We make a distinction here between spatially organized interfaces, where
multiple targets are displayed at the same time at different spatial locations,
and temporally organized interfaces, like Rapid Serial Visual Presentation,
where targets or small sets of targets are shown consecutively.
Traditionally, spatially organized interfaces have a higher information transfer
rate, but are also more gaze-dependent than temporally organized interfaces.
In such an interface, stimuli can be presented in a standard BCI paradigm, but
visuospatial attention can be decoded separately from the position of gaze,
which do not necessarily need to coincide.

This leads us to adopt a specific approach: improve the performance of a
spatially organized visual oddball ERP-based brain
computer interface by using a suited decoding strategy.
Hence, our working hypothesis is as follows:
\begin{quote}
	\textit{A spatially organized visual oddball ERP BCI with a suited decoder
		can achieve similar or higher information transfer rate in patient
		populations with eye motor deficits than other
		gaze-independent alternatives described in literature.}
\end{quote}
\todo{tone this down a bit to what can be proven}

To achieve our goal, we innovate on decoder development and interface design,
and we collect data from healthy control subjects, which allow us to test our decoders.
Finally, the findings of these experiments will be leveraged to develop a BCI
for patient use, which will be tested in a case-based patient study in which
both performance and user comfort will be evaluated.

\todo{\emph{Goal:} Enable communication for eye-motor impaired patients}
\todo{\emph{Method:} Design a comfortable interface that allows them to maximally exploit
their residual gaze capabilities, by leveraging a non-invasive high-ITR
visuospatial ERP-based BCI and improving ERP decoding performance (in general
and specifically in gaze-independent settings).}

\subsection{Approach}

\subsubsection{Interface design}

\todo{Overt, covert, split, free, illustration from poster}
\todo{clear definitions}

\subsubsection{Decoders}

During this PhD, we explored different lines in decoding strategies,
trying to tackle several problems that arise from gaze-independence, such as
the lack or decrease in amplitude of specific ERP components, and the increased
non-stationarity of the signal.
As mentioned, state-of-the-art decoders have poor performance in covert attention
settings.
The general goal is thus to design a machine learning classifier that represents
the ERP signal in a specific way, such that it becomes more robust to the problems
occurring in covert attention conditions.

\todo{Regularized spatiotemporal beamforming}
Due to the lack of the N2 component and decrease of amplitude of the
P3 component in covert attention
settings~\cite{Treder2010}, the signal-to-noise ratio (SNR) of
the ERP is lower than in overt attention settings.
Therefore, a straightforward way to reach satisfactory
gaze-independent decoding performance, might be by increasing overall ERP
decoding performance. A more accurate classifier could yield relative
improvements in those settings where performance is not yet near the achievable
maximum.

Therefore, we have improved upon an in-house developed, state-of-the art ERP
decoder, the spatiotemporal beamformer~\cite{Wittevrongel2016}, by reformulating
this classifier as a linear discrimination problem and
imposing regularizing constraints by structuring the noise covariance matrix
(STBF-struct).
Furthermore, these regularizing constraints impose temporal stationarity on
the background noise, which is of use in our next efforts to cope with the
non-stationarity of the P3 signal component.

\todo{Classifier-based Latency Estimation with Woody iterations}
Literature shows that better covert attention performance can be achieved by
counteracting the non-stationarity of the P3 ERP component in covert
attention settings~\cite{Arico2014}.
This can be done by estimating the latency of each single-trial ERP and aligning
the P3 peaks to all fall at the same moment.
The resulting aligned data and the set of latencies can then be used to train a
classifier that is more robust to jitter.

Existing latency estimation methods are either not applicable to the
classification problem of labeling unseen data, or are not robust enough to
deal with the low SNR of the ERP.
Classifier-based Latency Estimation (CBLE)~\cite{Mowla2017} is a technique that can leverage
ERP latency estimation in a decoding setting, but our results show that it
yields no improvement in gaze-independent settings.
We improved upon this technique and extended it to a probabilistic, iterative
method named Classifier-based Latency Estimation with Woody iterations (wCBLE).

\todo{Tensor discriminant analysis}

\subsubsection{Experiments}
\todo{Healthy participant study}
We designed a visual oddball interface that helps us study and adapt
to the effects of eye motor impairment on the ERP, when using existing state-of-the-art
decoders and on our own proposed decoders.
In a first series of experiments, we recorded data with this interface from healthy participants.
The goal of these experiments is to benchmark
gaze-independent ERP decoding algorithms.
We denote this dataset as the Covert Visuospatial Attention ERP dataset
(CVSA-ERP)
This study was approved by the Ethics Commission of University Hospital Leuven
(S62547).

Using a hexagonal lay-out interface, similar to the visual Hex-o-Spell proposed
by Treder \& Blankertz~\cite{Treder2010}, we present six flashing targets (without letters or
symbols) to the participant while the EEG and electro-oculogram (EOG) are
recorded, as well as eye gaze using eye tracking.
To simulate the dissociation between the eye gaze (visual attention) and the
intended target (mental attention) which could occur in patients with hampered eye
motor control, healthy participants are cued to operate the BCI in specific
visuospatial attention (VSA) conditions.
Each subject performs 5 different VSA conditions as illustrated in
Figure~\ref{fig:vsa-conditions}.
\todo{do we already want this here?}
\begin{figure}
	%\includegraphics[width=\linewidth]{figures/attention_modes.eps}
	\caption{Visuospatial attention (VSA) conditions in our experiment. The user is
		asked to dissociate their visual and mental attention by gazing at a
		crosshair while attending a target, depending on the prompted condition.}%
	\label{fig:vsa-conditions}
\end{figure}
\todo{new picture that shows more clearly a user operating freely where they
can gaze}
In the overt case, subjects are instructed to gaze at the cued
target they are also mentally attending; in the covert case, subjects are instructed
to gaze at the center of the screen while mentally attending to the cued target.
Finally, we introduce a VSA condition that is understudied in the context of
gaze-independent BCI development: split VSA.
In split VSA, the participant mentally focuses on one cued target while gazing at
another (the distractor).
Depending on the inter-target distance along the hull of the hexagon between the
attended target and the distractor we discern three split VSA subconditions:
the distractor is either clockwise or counterclockwise directly next to the
attended target ($d=1$), there is one other target between the attended target and
the distractor ($d=2$), or the distractor is opposite the intended target
($d=3$).

After data collection from healthy participants and decoder development, we
will carry out a patient study using a similar interface.
The goal of the patient study is to evaluate the influence of eye motor
deficits in patient populations on existing state-of-the-art decoders and on
our proposed decoding methods.
\todo{update with current results}

\todo{Patient study}
\todo{refer to chapters}
